---
title: "Overview of retrieved datasets"
output:
  pdf_document: default
  html_notebook: default
---

**Script description:** This script manipulates the raw data of the retrieved datasets to show summary tables and figures.

**Script structure:**

1.  Number of datasets and relevance categories.2. Queries performance - 3. Temporal duration - 4. Spatial range - 5. Temporal duration, spatial range, and relevance - 6. EBV data types - 7. Data format - 8. Datasets accessibility and relevance per year

(TODO)

9.  Journals

10. Taxonomic coverage

11. Dataset location

12. Temporal and spatial information location

```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
library(tidyr)
source("Functions.R")
```

Some parameters for the plots:

```{r}
my.theme<-theme(axis.text=element_text(size=15),
        axis.title = element_text(size = 17),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)),
        panel.background = element_rect(fill = 'white'))

```

#### Read data

Eliminate empty rows and those where dataset_relevance is left in blank (they correspond to referred publications in the retrieved datasets and are not used in these analyses):

```{r}

dataset <- read.csv("091222_Merged_datasets.csv", header = TRUE)

dataset <- dataset[dataset$dataset_relevance != "",]

dataset$id_query

```

## 1. Number of datasets and relevance categories

```{r}
total_publications <- length(dataset$dataset_relevance)

print(paste("A total of", total_publications, "publications retrieved"))
  
```

```{r}
dataset$dataset_relevance <- as.factor(dataset$dataset_relevance)

df_N_relevance <- count_by_relevance(dataset)

df_N_relevance
```

## 2. Queries performance

### 2.1. N publications and relevance categories per query

#### Grouped queries

```{r}

dataset$id_query <- as.factor(dataset$id_query)

df_queries_counts <- compute_df_n_relevance_queries(data = dataset)

df_queries_counts
```

Plot the results:

```{r}
plot_queries_relevance <- plot_queries_relevance_raw(df_queries_counts)

plot_queries_relevance
```

-   Query 4 delivers the highest number of relevant datasets (highest number of High, Moderate and Low relevance datasets), but also retrieves by far the higher amount of publication.

-   Some queries do not add any relevant publication (e.g. "0,6,5,7" or "7,2,9").


#### Individual queries


```{r}

#split rows in function of character comas

dataset_q <- dataset %>%               
  separate_rows(id_query, sep=",") 


dataset_q$id_query  <- as.factor(dataset_q$id_query)


df_queries_counts_ind <- compute_df_n_relevance_queries(data = dataset_q)

df_queries_counts_ind
```

Plot the results

```{r}
plot_queries_relevance_ind <- plot_queries_relevance_raw(df_queries_counts_ind)

plot_queries_relevance_ind
```

- query 4 shows the highest number of relevant datasets for high, moderate and low relevance datasets, but retrieves almost double number of articles (~70) compared to the rest of queries.

- queries 5 and 7 are able to retrieve some relevant datasets with ~25 publications retrieved.


### 2.2 Classify queries by their relevance (index)

Puctuations for the relevance categories:

-   High = 5
-   Mod = 2
-   Low = 0.5
-   non_relevant = 0

*Index 1*: is the result of summing the puntuations assigned to each relevance category

Index 1 = sum punctuations = (N High \* 5) + (N Mod \* 2) + (N Low \* 0.5)

*Index 2*: is the Index1 but accounting for the total number of publications retrieved by the query.

Index 2 = sum punctuations / n publications retrieved


#### Grouped queries

```{r}

df_queries_counts_index <- compute_index12(df_counts = df_queries_counts, order_based_index = "2") # "1" for ordering according to index 1, "2" for index 2

df_queries_counts_index

```

Queries 3, 4, 0, and 6 are the top 4 index1 values. Queries 3 and 4 have far higher index than the rest.

Plot the results:

```{r}

plot_queries_index12 <- plot_queries_index(df = df_queries_counts_index)

plot_queries_index12
```

-   Index 1 shows that queries "3" and "4" offer the higher perfomance, followed by "6" and "5".

-   Index 2 shows that querie "6,7,5" offer the highest performance, followed by "7,8", "6,7", and "5,6,7,8,9"


#### individual queries

```{r}
df_queries_counts_index_ind <- compute_index12(df_counts = df_queries_counts_ind, order_based_index = "2") # "1" for ordering according to index 1, "2" for index 2

df_queries_counts_index_ind
```


Plot the results

```{r}
plot_queries_index12_ind <- plot_queries_index(df = df_queries_counts_index_ind)

plot_queries_index12_ind
```



## 3. Temporal duration

### 3.1 Temporal duration counts

How many publications without temporal duration data?

```{r}
nNa <- count_not.reported_temporal.duration(dataset)

print(paste(nNa, "publications without temporal duration data"))

```

Publication counts by temporal duration (years):

```{r}
  
df_duration_counts <- count_durations(df = dataset, order_by = "counts")

df_duration_counts
```

Plot the results:

```{r}
plot_temp_duration <- plot_duration_counts(df = df_duration_counts, 
                                           counts_Na = nNa)

plot_temp_duration
```

Average duration (for those with data):

```{r}

dataset_temp_duration <- subset(dataset, 
                                  temporal_duration_y > 0 | temporal_duration_y == "no",
                                  select = c(temporal_duration_y,id_query))
  
  dataset_temp_duration$temporal_duration_y <- as.numeric(dataset_temp_duration$temporal_duration_y)
  
  

print(paste("mean of",
            round(mean(na.omit(dataset_temp_duration$temporal_duration_y)),digits = 1),
            "years"))
```

### 3.1 Temporal duration counts per relevance category

```{r}


plot_duration_relevance.categories <- plot_duration_relevance(df = dataset,
                                                              counts_Na = nNa)

plot_duration_relevance.categories
```

## 4. Spatial range

### 4.1 count spatial range publications

WARNING: publications that cant be accessed are not counted

How many publications without spatial range data?

```{r}

n_not_reported <- count_not.reported_spatial_range(dataset)

print(paste(n_not_reported, "publications without spatial range data"))

```

Spatial ranges are divided according to the thresholds established to determine a low, moredate and high spatial range: \<5000, 500-15000, \>15000

```{r}

plot_spatial_range_counts <- plot_spat.range_counts(dataset)

plot_spatial_range_counts
```

Average spatial range (for those with data):

```{r}

dataset1 <- dataset[dataset$dataset_relevance != "cant access",]
  
  dataset1$dataset_relevance <- as.factor(dataset1$dataset_relevance)
  
  spatial_range_km2_vec <- dataset1$spatial_range_km2[dataset1$spatial_range_km2 != ""]
  
  spatial_range_km2_vec <- spatial_range_km2_vec[!is.na(spatial_range_km2_vec)]
  
  spatial_range_km2_vec <- as.numeric(spatial_range_km2_vec) 


print(paste("mean of",
            round(mean(na.omit(spatial_range_km2_vec)), digits = 0),
            "km2"))
```

## 5. Temporal duration, spatial range, relevance

```{r}

plot_spatial_temporal_relevance <- plot_spat_temp_relevance(df = dataset)

plot_spatial_temporal_relevance
```

## 6. EBV data types

```{r}

df_data_type_counts <- compute_df_data.type(df = dataset)

df_data_type_counts

```

Plot the results

```{r}

plot_data_type <- plot_data.type_counts(df_data_type_counts)

plot_data_type

```

## 7. Data format

```{r}

plot_data_format_counts <- plot_data.type_format(dataset)

plot_data_format_counts

```

## 8. Datasets accessibility and relevance per year

Raw plot relevance per year

```{r}

plot_relevance_access_year <- plot_relevance_year(dataset)

plot_relevance_access_year

```

Relevance per year (intervals):

```{r}

df_relevance_year.range <- compute_df_relevance_year.range(dataset)

df_relevance_year.range

```

Plot the results:

```{r}

plot_relevance_year.range <- plot_relevance_year.range(dataset)

plot_relevance_year.range

```




## 9. Journals


Count publications per journal

```{r}

dataset.j <- dataset[dataset$Journal != "" & dataset$Journal != "no",]

df_journals_counts <- as.data.frame(table(dataset.j$Journal))

colnames(df_journals_counts) = c("Journal", "Publication_counts")

df_journals_counts <- df_journals_counts[order(-df_journals_counts$Publication_counts),]


df_journals_counts

```

```{r}

print(paste("A total of", length(df_journals_counts$Journal), "journals" ))

```

```{r}

```


## 10. Location of information


