---
title: "R Notebook"
output: html_notebook
---

**Script description:** This script manipulates the raw data of the retrieved datasets to show summary tables and figures.


```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
library(tidyr)
library(stringr)
library(ggrepel)
library(readxl)

source("Functions.R")
```

Some parameters for the plots:

```{r}
my.theme<-theme(axis.text=element_text(size=15),
        axis.title = element_text(size = 17),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)),
        panel.background = element_rect(fill = 'white'))

```




#### Read data


```{r}

dataset <- read_excel("df_repository.xlsx")
table(dataset$valid_yn)

```


```{r}

table(dataset$reason_non_valid[which(dataset$valid_yn=="no")])
```


Eliminate empty rows and those where dataset_relevance is left in blank (they correspond to referred publications in the retrieved datasets and are not used in these analyses):

```{r}
dataset <- dataset[dataset$dataset_relevance != "",]
dataset <- dataset %>% dplyr::filter(valid_yn == "yes")
dataset$dataset_relevance <- as.factor(dataset$dataset_relevance)

dataset <- dataset[dataset$source != "semantic_scholar",]
dataset$id_query  <- stringr::str_trim(dataset$id_query )
dataset$dataset_relevance  <- str_trim(dataset$dataset_relevance )
head(dataset)
```



## 1. Number of datasets and relevance categories


```{r}

df_N_relevance <- count_by_relevance(dataset)

df_N_relevance
```







## 2. Queries performance

### 2.1. N publications and relevance categories per query


#### Individual queries

```{r}

convert_query = data.frame("id_query" = as.character(seq(0,10)), "query" = c(
"survey + species",
"time series + species",
"inventory + species",
"species",
"abundance + species",
"occurrence + species",
"population + species",
"sites + species",
"sampling + species",
"collection + species",
"density + species"
  
))


```

```{r}

#split rows in function of character comas

dataset_q <- dataset %>%               
  separate_rows(id_query, sep=",") 
dataset_q <- merge(dataset_q, convert_query, by.all="id_query")

length(dataset_q$id_query == "species")

dataset_q$query  <- factor(dataset_q$query,
                              levels= c("species", "occurrence + species", "inventory + species", "collection + species", "sampling + species", "survey + species" , "population + species", "sites + species", "density + species", "abundance + species" , "time series + species" ))





```




```{r}
df_queries_counts_ind <- dataset_q %>% 
  group_by(dataset_relevance, query) %>% 
  summarise(n = n()) 

df_queries_counts_ind <- df_queries_counts_ind %>% dplyr::filter(dataset_relevance %in% c("X", "L", "M", "H")) %>% mutate(dataset_relevance = factor(dataset_relevance))



plot_queries_relevance_counts <- ggplot(df_queries_counts_ind, 
                                          aes(x=query, y=n, group = dataset_relevance, 
                                              color = dataset_relevance,shape=dataset_relevance )) +
  geom_segment( aes(x=query ,xend=query, y=0, yend=max(n)), color="grey") +
    geom_point(size=4, alpha = 0.8) +
    coord_flip() +
    # theme_ipsum() +
    theme(
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_blank(),
      legend.position="top"
    ) +
    xlab("Query ID")+
    ylab("publications counts") +
    scale_color_manual(name = "dataset relevance",
                       values = c( "red","purple2", "dodgerblue", "black"))+
    scale_shape_manual(name = "dataset relevance", 
                       values = c(19, 19, 19, 1))+
    my.theme

plot_queries_relevance_counts

ggsave("plots_repo/plot_queries_raw.png", height = 5, width = 9)

```


### 2.2 Classify queries by their relevance (index)

Puctuations for the relevance categories:

-   High = 5
-   Mod = 2
-   Low = 0.5
-   non_relevant = 0

*Index 1*: is the result of summing the puntuations assigned to each relevance category

Index 1 = sum punctuations = (N High \* 5) + (N Mod \* 2) + (N Low \* 0.5)

*Index 2*: is the Index1 but accounting for the total number of publications retrieved by the query.

Index 2 = sum punctuations / n publications retrieved


#### Grouped queries

```{r}

df_queries_counts_index <- compute_index12(df_counts = df_queries_counts, order_based_index = "2") # "1" for ordering according to index 1, "2" for index 2

df_queries_counts_index

```

Queries 3, 4, 0, and 6 are the top 4 index1 values. Queries 3 and 4 have far higher index than the rest.

Plot the results:

```{r}

plot_group.queries_index12 <- plot_queries_index(df = df_queries_counts_index)

plot_group.queries_index12

#ggsave("plots_repo/plot_group.queries_index12.png", height = 5, width = 11)
```

-   Index 1 shows that queries "3" and "4" offer the higher perfomance, followed by "6" and "5".

-   Index 2 shows that querie "6,7,5" offer the highest performance, followed by "7,8", "6,7", and "5,6,7,8,9"


#### individual queries

```{r}
df_queries_counts_index_ind <- compute_index12(df_counts = df_queries_counts_ind, order_based_index = "2") # "1" for ordering according to index 1, "2" for index 2

df_queries_counts_index_ind
```


Plot the results

```{r}
plot_queries_index12_ind <- plot_queries_index(df = df_queries_counts_index_ind)

plot_queries_index12_ind

#ggsave("plots_repo/plot_ind.queries_index12.png", height = 5, width = 11)
```



### 2.3. F SCORE

Precision
Recall

Binary view: M,H relevant | L,X unrevelant


N relevant datasets -> True Positives (TP)

N unrelevant datasets -> False Positives (FP)

Sum of all relevant publications of the dataset that are not detected = FN

PRECISION = TP / TP + FP

RECALL = TP / TP + FN


F SCORE = 2* precision * recall / precision + recall



Split grouped queries into individual ones


```{r}
dataset_q
```





```{r}

df_zscores_queries <- calculate_z.score_queries(df = dataset)

df_zscores_queries$Fscore <- round(df_zscores_queries$Fscore, 2)
df_zscores_queries$Precision <- round(df_zscores_queries$Precision, 2)
df_zscores_queries$Recall <- round(df_zscores_queries$Recall, 2)


df_zscores_queries

write.table(df_zscores_queries, file = "df_zscores_queries.txt", sep = ",", quote = FALSE, row.names = F)

```

Plot results

```{r}

mid = mean(df_zscores_queries$Fscore)

plot_queries_scores <- ggplot(df_zscores_queries, aes(x = Precision, y = Recall, colour = Fscore)) +
  geom_point(size = 10)+
  geom_label_repel(aes(label = query),alpha = 0.75, 
                   label.padding=.1)+
  scale_color_gradient2(midpoint=mid, low="dodgerblue", mid="purple2",
                     high="red", space ="Lab" )+
  theme_bw()+
  my.theme

plot_queries_scores

#ggsave("plots_repo/plot_queries_scores.png", height = 7, width = 7)

```

```{r}

combined_plot_queries <- ggarrange(plot_queries_relevance_counts,
                                   plot_queries_scores, 
                                   nrow = 1,
                                   ncol = 2)

ggsave("plots_repo/combined_plot_queries.png",
       height = 6,
       width = 13)

```



## 3. Temporal range and duration


## Datasets temporal range





```{r}



dataset_temp_range <- dataset[,c("url", "temp_range_i", "temp_range_f", "time_series")]

dataset_temp_range <- dataset_temp_range[-which(is.na(dataset_temp_range$temp_range_i)),]

dataset_temp_range <- as.data.frame((dataset_temp_range))

#dataset_temp_range <- melt(dataset_temp_range, measure.vars = c("temp_range_i", "temp_range_f"))

#dataset_temp_range <- melt(dataset_temp_range, id ="url")

dataset_temp_range$temp_range_i <- as.numeric(dataset_temp_range$temp_range_i)
dataset_temp_range$temp_range_f <- as.numeric(dataset_temp_range$temp_range_f)

dataset_temp_range <- dataset_temp_range[-which(dataset_temp_range$temp_range_i == -20000),]

dataset_temp_range[which(dataset_temp_range$temp_range_i == dataset_temp_range$temp_range_f),"temp_range_f"] <- dataset_temp_range[which(dataset_temp_range$temp_range_i == dataset_temp_range$temp_range_f),"temp_range_f"]+0.5

dataset_temp_range$temp_range_i[which(dataset_temp_range$temp_range_i < 1930)] <- 1930

p_temp_range <- ggplot(dataset_temp_range, aes(y=url, color = time_series)) +
  geom_segment(aes(x=temp_range_i, xend=temp_range_f, y=url, yend=url), size=3)+
  xlab("year") +
  ylab ("datasets")+
  theme_bw() +
  my.theme +
  theme(axis.text.y=element_blank())+
  scale_color_manual(values = c("gray37", "#F8766D"))+
  theme(legend.position = "none") +
  xlim(1930, round(max(dataset_temp_range$temp_range_f),0))+
  geom_label(
    label="1875", 
    x=1930,
    y=5,
    #label.padding = unit(0.2, "lines"), # Rectangle size around label
    label.size = 0.1,
    color = "black"
   # fill="white"
  )

p_temp_range



#ggsave("plots_repo/temporal_range.png", height = 7, width = 7)

```

```{r}
min(dataset_temp_range$temp_range_i)

dataset_temp_range[dataset_temp_range$temp_range_i == 1875,]
```


### 3.1 Temporal duration counts

How many publications without temporal duration data?

```{r}


#nNa <- count_not.reported_temporal.duration(dataset)

print(paste(length(which(dataset$temporal_duration_position == "no")), "publications without temporal duration data"))


```


Publication counts by temporal duration (years):

```{r}
  
df_duration_counts <- count_durations(df = dataset, order_by = "counts")

df_duration_counts
```

Plot the results:

```{r}
plot_temp_duration <- plot_duration_counts(df = df_duration_counts, 
                                           counts_Na = nNa)

plot_temp_duration

ggsave("plots_repo/plot_temp_duration.png", height = 5, width = 9)

```

Average duration (for those with data):

```{r}

dataset_temp_duration <- subset(dataset, 
                                  temporal_duration_y > 0 | temporal_duration_y == "no",
                                  select = c(temporal_duration_y,id_query))
  
  dataset_temp_duration$temporal_duration_y <- as.numeric(dataset_temp_duration$temporal_duration_y)
  

print(paste("mean of",
            round(mean(na.omit(dataset_temp_duration$temporal_duration_y)),digits = 1),
            "years"))
```



## 4. Spatial range

### 4.1 count spatial range publications

WARNING: publications that cant be accessed are not counted

How many publications without spatial range data?




```{r}

n_not_reported <- count_not.reported_spatial_range(dataset)

print(paste(n_not_reported, "publications without spatial range data"))

```

Spatial ranges are divided according to the thresholds established to determine a low, moredate and high spatial range: \<5000, 500-15000, \>15000

```{r}

plot_spatial_range_counts <- plot_spat.range_counts(dataset)


#ggsave("plots_repo/plot_spatial_range_counts.png", height = 6, width = 6)



```

Average spatial range (for those with data):

```{r}

dataset1 <- dataset[dataset$dataset_relevance != "cant access",]
  
  dataset1$dataset_relevance <- as.factor(dataset1$dataset_relevance)
  
  spatial_range_km2_vec <- dataset1$spatial_range_km2[dataset1$spatial_range_km2 != ""]
  
  spatial_range_km2_vec <- spatial_range_km2_vec[!is.na(spatial_range_km2_vec)]
  
  spatial_range_km2_vec <- as.numeric(spatial_range_km2_vec) 


print(paste("mean of",
            round(mean(na.omit(spatial_range_km2_vec)), digits = 0),
            "km2"))
```

## 5. Temporal duration, spatial range, relevance

```{r}

plot_spatial_temporal_relevance <- plot_spat_temp_relevance(df = dataset)


plot_spatial_temporal_relevance

#ggsave("plots_repo/plot_spatial_temporal_relevance.png", height = 6, width = 9)

```

## 6. EBV data types

```{r}

df_data_type_counts <- compute_df_data.type(df = dataset)

df_data_type_counts


```

Plot the results

```{r}


plot_data_type <- plot_data.type_counts(df_data_type_counts)

plot_data_type

#ggsave("plots_repo/plot_data_type.png", height = 6, width = 9)



```

### Combine plots

V1

```{r}

plot_combined_corpus <- ggarrange(
plot_temp_duration,
plot_spatial_range_counts,
plot_spatial_temporal_relevance,
plot_data_type,
nrow = 2,
ncol = 2,
  labels = c("A", "B", "C","D"),
heights = c(1,1.22))

plot_combined_corpus

ggsave("plots_repo/plot_combined_corpus.png", height = 10, width = 12)
```

V2

```{r}

plot_combined_corpus.v2 <- ggarrange(
p_temp_range,
plot_spatial_range_counts,
plot_spatial_temporal_relevance,
plot_data_type,
nrow = 2,
ncol = 2,
  labels = c("A", "B", "C","D"),
heights = c(1,1.22))

plot_combined_corpus.v2

ggsave("plots_repo/plot_combined_corpus.v2.png", height = 10, width = 12)
```

## 7. Data format


Those datasets in the repository that are relevance category X don't have format information.

```{r}

#plot_data_format_counts <- plot_data.type_format(dataset)

#plot_data_format_counts

#ggsave("plots_repo/plot_data_format_counts.png", height = 7, width = 9)


```






## 9. Journals


Count publications per journal

```{r}

dataset.j <- dataset[dataset$journal != "" & dataset$journal != "no",]

df_journals_counts <- as.data.frame(table(dataset.j$journal))

colnames(df_journals_counts) = c("Journal", "Publication_counts")

df_journals_counts <- df_journals_counts[order(-df_journals_counts$Publication_counts),]


df_journals_counts


```




```{r}

print(paste("A total of", length(df_journals_counts$Journal), "journals" ))

```


## Location information - temporal range


```{r}

dataset_repos <- read_excel("df_repository.xlsx")

dataset_repos <- dataset_repos[which(dataset_repos$title != ""),]
dataset_repos <- dataset_repos %>% dplyr::filter(valid_yn == "yes")

tempr_position <- dataset_repos$temporal_range_position[-which(is.na(dataset_repos$temporal_range_position))]

tempr_position  <- unlist(strsplit(tempr_position,","))

tempr_position[which(tempr_position == " dataset")] <- gsub(" ", "", tempr_position[which(tempr_position == " dataset")])

tempr_position[which(tempr_position == " source publication text")] <- "source publication text"

tempr_position[which(tempr_position == " source link")] <- "source link"

tempr_position



```


```{r}


tempd_position <- dataset_repos$temporal_duration_position[-which(is.na(dataset_repos$temporal_duration_position))]
tempd_position  <- unlist(strsplit(tempd_position,","))

tempd_position

```


```{r}

spatr_position <- dataset_repos$spatial_range_position[-which(is.na(dataset_repos$spatial_range_position))]
spatr_position  <- unlist(strsplit(spatr_position,","))

spatr_position

```


```{r}

feature <- rep("temporal range position", times = length(tempr_position))

df_tempr <- data.frame(tempr_position, feature)

feature <- rep("temporal duration position", times = length(tempd_position))

df_tempd <- data.frame(tempd_position, feature)

feature <- rep("Spatial range position", times = length(spatr_position))

df_spatr <- data.frame(spatr_position, feature)

colnames(df_tempr) <- c("position", "feature")
colnames(df_tempd) <- c("position", "feature")
colnames(df_spatr) <- c("position", "feature")

df_positions <- rbind(df_tempr,df_tempd,df_spatr)

```


```{r}

levels(as.factor(df_positions$position))

df_positions$position[which(df_positions$position == " source publication text")] <- "source publication text"
df_positions$position[which(df_positions$position == "no")] <- "not given"

df_positions$position <- factor(df_positions$position, levels=c("not given", "source publication text", "dataset", "source link", "source link abstract"))

df_positions <- as.data.frame(table(df_positions))

ggplot(df_positions, aes(x=position, fill = feature)) +
geom_bar(position = "dodge") +  labs(x='Team')+
  theme_bw() +
  my.theme+
  theme(axis.text.x = element_text(angle = 0, hjust=0.45,vjust=0.2))+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))
  #scale_fill_manual(values = c(""))


levels(df_positions$feature)

ggplot(df_positions, aes(x=position, y = Freq, group = feature, color = feature)) +
   geom_segment( aes(x=position ,xend=position, y=0, yend=max(Freq)), color="grey") +
    geom_point(size=4, alpha = 0.8) +
    coord_flip() +
  theme_bw() +
  my.theme+
  scale_color_manual(values = c("deepskyblue2","firebrick1","indianred4"))+
  scale_shape_manual(values = c(11,2,3))+
  ylab("publication counts")

  
ggsave("plots_repo/location_spat.temp.features_repos.png", height = 5, width = 9)


```



## Spatial information format


```{r}

dataset_repos$geospatial_info_article_text
dataset_repos$geospatial_info_dataset
dataset_repos$geospatial_info_repo_page_text





geospatial_info_article_text <- dataset_repos$geospatial_info_article_text[-which(is.na(dataset_repos$geospatial_info_article_text))]

geospatial_info_dataset <- dataset_repos$geospatial_info_dataset[-which(is.na(dataset_repos$geospatial_info_dataset))]

geospatial_info_repo_page_text <- dataset_repos$geospatial_info_repo_page_text[-which(is.na(dataset_repos$geospatial_info_repo_page_text))]






geospatial_info_article_text  <- unlist(strsplit(geospatial_info_article_text,","))

geospatial_info_dataset  <- unlist(strsplit(geospatial_info_dataset,","))

geospatial_info_repo_page_text  <- unlist(strsplit(geospatial_info_repo_page_text,","))


geospatial_info_article_text[which(geospatial_info_article_text == " administrative unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == " range coordites")] <- "range coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == " site coordites")] <- "site coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == "site coordites")] <- "site coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == " geological feature me")] <- "geographic feature"
geospatial_info_article_text[which(geospatial_info_article_text == "geological feature me")] <- "geographic feature"
geospatial_info_article_text[which(geospatial_info_article_text == " Map")] <- "Map"
geospatial_info_article_text[which(geospatial_info_article_text == "Map")] <- "map"
geospatial_info_article_text[which(geospatial_info_article_text == "administration unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == "Administrative unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == "no")] <- "not given"


geospatial_info_dataset[which(geospatial_info_dataset == " site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "  site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "sample coordites")] <- "sample coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == " sample coordites")] <- "sample coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == " administrative unit")] <- "administrative unit"
geospatial_info_dataset[which(geospatial_info_dataset == " site IDs")] <- "site IDs"
geospatial_info_dataset[which(geospatial_info_dataset == " IDs")] <- "site IDs"
geospatial_info_dataset[which(geospatial_info_dataset == " geographic feature")] <- "geographic feature"
geospatial_info_dataset[which(geospatial_info_dataset == "no")] <- "not given"


geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "geological feature me")] <- "geographic feature"
geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "Administrative unit")] <- "administrative unit"
geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "no")] <- "not given"

```


```{r}

location <- rep("article text", times = length(geospatial_info_article_text))

df_article <- data.frame(geospatial_info_article_text, location)

location <- rep("repository page", times = length(geospatial_info_repo_page_text))

df_repo<- data.frame(geospatial_info_repo_page_text, location)

location <- rep("dataset", times = length(geospatial_info_dataset))

df_dataset <- data.frame(geospatial_info_dataset, location)

colnames(df_article) <- c("information", "location")
colnames(df_repo) <- c("information", "location")
colnames(df_dataset) <- c("information", "location")

df_locations <- rbind(df_article,df_repo,df_dataset)

df_locations$information[which(df_locations$information == "site IDs") ] <- "sites IDs"
df_locations$information[which(df_locations$information == "site coordites") ] <- "site coordinates"
df_locations$information[which(df_locations$information == "Map") ] <- "map"
df_locations$information[which(df_locations$information == "geologial feature") ] <- "geographic feature"
df_locations$information[which(df_locations$information == "geographical feature") ] <- "geographic feature"

```


```{r}

df_locations <- as.data.frame(table(df_locations))


df_locations$location <- factor(df_locations$location, levels=c("article text", "repository page", "dataset"))



ggplot(df_locations, aes(x=information, y = Freq, group = location, color = location)) +
   geom_segment( aes(x=information ,xend=information, y=0, yend=max(Freq)), color="grey") +
    geom_point(size=4, alpha = 0.5) +
    coord_flip() +
  theme_bw() +
  my.theme+
 scale_color_manual(values = c("blue4","red1","green4"))+
 # scale_shape_manual(values = c(11,2,3))+
  ylab("publication counts")

  
ggsave("plots_repo/location_type_geoinfo.png", height = 5, width = 9)


```