---
title: "R Notebook"
output: html_notebook
---

**Script description:** This script manipulates the raw data of the retrieved datasets to show summary tables and figures.


```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(hrbrthemes)
library(ggpubr)
library(tidyr)
library(stringr)
library(ggrepel)
library(readxl)

source("Functions.R")
```

Some parameters for the plots:

```{r}
my.theme<-theme(axis.text=element_text(size=15),
        axis.title = element_text(size = 17),
        legend.text=element_text(size=10),
        legend.title = element_text(size=12),
        plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)),
        panel.background = element_rect(fill = 'white'))

```




#### Read data


```{r}

dataset <- read_excel("df_repository.xlsx")

```

Count Valid/non-valid datasets:

```{r}

table(dataset$valid_yn)

```


count reasons for non-valid assignment:


```{r}

table(dataset$reason_non_valid[which(dataset$valid_yn=="no")])
```


Eliminate empty rows and those where dataset_relevance is left in blank (they correspond to referred publications in the retrieved datasets and are not used in these analyses):

```{r}
dataset <- dataset[dataset$dataset_relevance != "",]
dataset <- dataset %>% dplyr::filter(valid_yn == "yes")
dataset$dataset_relevance <- as.factor(dataset$dataset_relevance)

dataset <- dataset[dataset$source != "semantic_scholar",]
dataset$id_query  <- stringr::str_trim(dataset$id_query )
dataset$dataset_relevance  <- str_trim(dataset$dataset_relevance )

head(dataset)
```



## 1. Number of datasets and relevance categories


```{r}

df_N_relevance <- count_by_relevance(dataset)

df_N_relevance
```




## 2. Queries performance

### 2.1. N publications and relevance categories per query


```{r}

convert_query = data.frame("id_query" = as.character(seq(0,10)), "query" = c(
"survey + species",
"time series + species",
"inventory + species",
"species",
"abundance + species",
"occurrence + species",
"population + species",
"sites + species",
"sampling + species",
"collection + species",
"density + species"
  
))


```

```{r}

#split rows in function of character comas

dataset_q <- dataset %>%               
  separate_rows(id_query, sep=",") 


dataset_q <- merge(dataset_q, convert_query, by.all="id_query")


dataset_q$query  <- factor(dataset_q$query,
                           levels= c("species", "occurrence + species",
                                     "inventory + species", "collection + species",
                                     "sampling + species", "survey + species" , 
                                     "population + species", "sites + species", 
                                     "density + species", "abundance + species",
                                     "time series + species" ))


df_queries_counts_ind <- dataset_q %>% 
  group_by(dataset_relevance, query) %>% 
  summarise(n = n()) 

df_queries_counts_ind <- df_queries_counts_ind %>% 
  dplyr::filter(dataset_relevance %in% c("X", "L", "M", "H")) %>% 
  mutate(dataset_relevance = factor(dataset_relevance))


df_queries_counts_ind

```


Plot the results:

```{r}


plot_queries_relevance_counts <- ggplot(df_queries_counts_ind, 
                                          aes(x=query, y=n, group = dataset_relevance, 
                                              color = dataset_relevance,shape=dataset_relevance )) +
  geom_segment( aes(x=query ,xend=query, y=0, yend=max(n)), color="grey") +
    geom_point(size=4, alpha = 0.8) +
    coord_flip() +
    # theme_ipsum() +
    theme(
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_blank(),
      legend.position="top"
    ) +
    xlab("Query ID")+
    ylab("publications counts") +
    scale_color_manual(name = "dataset relevance",
                       values = c( "red","purple2", "dodgerblue", "black"))+
    scale_shape_manual(name = "dataset relevance", 
                       values = c(19, 19, 19, 1))+
    my.theme


#ggsave("plots_repo/plot_queries_raw.png", height = 5, width = 9)

```



### 2.2. F SCORE

- F SCORE = 2* precision * recall / precision + recall

- PRECISION = TP / TP + FP

- RECALL = TP / TP + FN

N relevant datasets -> True Positives (TP)

N unrelevant datasets -> False Positives (FP)

Sum of all relevant publications of the dataset that are not detected = FN


- Binary view: M,H relevant | L,X unrevelant




```{r}

df_zscores_queries <- calculate_z.score_queries(df = dataset)

df_zscores_queries$Fscore <- round(df_zscores_queries$Fscore, 2)
df_zscores_queries$Precision <- round(df_zscores_queries$Precision, 2)
df_zscores_queries$Recall <- round(df_zscores_queries$Recall, 2)


df_zscores_queries

#write.table(df_zscores_queries, file = "df_zscores_queries.txt", sep = ",", quote = FALSE, row.names = F)

```

Plot results:

```{r}

mid = mean(df_zscores_queries$Fscore)

plot_queries_scores <- ggplot(df_zscores_queries, aes(x = Precision, y = Recall, colour = Fscore)) +
  geom_point(size = 10)+
  geom_label_repel(aes(label = query),alpha = 0.75, 
                   label.padding=.1)+
  scale_color_gradient2(midpoint=mid, low="dodgerblue", mid="purple2",
                     high="red", space ="Lab" )+
  theme_bw()+
  my.theme

#ggsave("plots_repo/plot_queries_scores.png", height = 7, width = 7)

```

```{r}

combined_plot_queries <- ggarrange(plot_queries_relevance_counts,
                                   plot_queries_scores, 
                                   nrow = 1,
                                   ncol = 2)

combined_plot_queries

#ggsave("plots_repo/combined_plot_queries.png",height = 6,width = 13)

```



## 3. Temporal range and duration

Discard datasets with negligible relevance for all the analyses from here - IMPORTANT TO NOTE - now the dataset size is n = 92 (discarting 16 negligible)

```{r}
dataset <- dataset[dataset$dataset_relevance != "X",]
```



### Temporal range




```{r}

# Select variables of interest

dataset_temp_range <- dataset[,c("url", "temp_range_i", "temp_range_f", "time_series")]

# Eliminate Nas

dataset_temp_range <- dataset_temp_range[-which(is.na(dataset_temp_range$temp_range_i)),]

dataset_temp_range <- as.data.frame((dataset_temp_range))

dataset_temp_range$temp_range_i <- as.numeric(dataset_temp_range$temp_range_i)
dataset_temp_range$temp_range_f <- as.numeric(dataset_temp_range$temp_range_f)

print(paste(nrow(dataset_temp_range), "with temporal range data"))


# Eliminate outlier

dataset_temp_range <- dataset_temp_range[-which(dataset_temp_range$temp_range_i == -20000),]

dataset_temp_range[which(dataset_temp_range$temp_range_i == dataset_temp_range$temp_range_f),"temp_range_f"] <- dataset_temp_range[which(dataset_temp_range$temp_range_i == dataset_temp_range$temp_range_f),"temp_range_f"]+0.5

# cut lower limit range for outlier

dataset_temp_range$temp_range_i[which(dataset_temp_range$temp_range_i < 1930)] <- 1930






```

Plot the results:

```{r}

p_temp_range <- ggplot(dataset_temp_range, aes(y=url, color = time_series)) +
  geom_segment(aes(x=temp_range_i, xend=temp_range_f, y=url, yend=url), linewidth=3)+
  xlab("year") +
  ylab ("datasets")+
  theme_bw() +
  my.theme +
  theme(axis.text.y=element_blank())+
  scale_color_manual(values = c("gray37", "#F8766D"))+
  theme(legend.position = "none") +
  xlim(1930, round(max(dataset_temp_range$temp_range_f),0))+
  geom_label(
    label="1875", 
    x=1930,
    y=5,
    #label.padding = unit(0.2, "lines"), # Rectangle size around label
    label.size = 0.1,
    color = "black"
   # fill="white"
  )

#ggsave("plots_repo/temporal_range.png", height = 7, width = 7)

```




### 3.1 Temporal duration counts

How many publications without temporal duration data?

```{r}

#nNa <- count_not.reported_temporal.duration(dataset)

print(paste(length(which(dataset$temporal_duration_position == "no")), "publications without temporal duration data"))


```


Publication counts by temporal duration (years):

```{r}
  
df_duration_counts <- count_durations(df = dataset, order_by = "counts")

df_duration_counts
```

Plot the results:

```{r}
plot_temp_duration <- plot_duration_counts(df = df_duration_counts, 
                                           counts_Na = nNa)

plot_temp_duration

#ggsave("plots_repo/plot_temp_duration.png", height = 5, width = 9)

```

Average duration (for those with data):

```{r}

dataset_temp_duration <- subset(dataset, 
                                  temporal_duration_y > 0 | temporal_duration_y == "no",
                                  select = c(temporal_duration_y,id_query))
  
  dataset_temp_duration$temporal_duration_y <- as.numeric(dataset_temp_duration$temporal_duration_y)
  

print(paste("mean of",
            round(mean(na.omit(dataset_temp_duration$temporal_duration_y)),digits = 1),
            "years"))
```



### 4. Spatial range

#### 4.1 count spatial range publications

Publications that cant be accessed are not counted.

How many publications without spatial range data?


```{r}

n_not_reported <- count_not.reported_spatial_range(dataset)

print(paste(n_not_reported, "publications without spatial range data"))

```

Average spatial range:

```{r}

dataset1 <- dataset[dataset$dataset_relevance != "cant access",]
  
  dataset1$dataset_relevance <- as.factor(dataset1$dataset_relevance)
  
  spatial_range_km2_vec <- dataset1$spatial_range_km2[dataset1$spatial_range_km2 != ""]
  
  spatial_range_km2_vec <- spatial_range_km2_vec[!is.na(spatial_range_km2_vec)]
  
  spatial_range_km2_vec <- as.numeric(spatial_range_km2_vec) 


print(paste("mean of",
            round(mean(na.omit(spatial_range_km2_vec)), digits = 0),
            "km2"))
```



Plot the results:
Spatial ranges are divided according to the thresholds established to determine a low, moredate and high spatial range: \<5000, 500-15000, \>15000

```{r}

plot_spatial_range_counts <- plot_spat.range_counts(dataset)


#ggsave("plots_repo/plot_spatial_range_counts.png", height = 6, width = 6)

```



## 5. Temporal duration, spatial range, relevance


```{r}

plot_spatial_temporal_relevance <- plot_spat_temp_relevance(df = dataset)


#ggsave("plots_repo/plot_spatial_temporal_relevance.png", height = 6, width = 9)

```

## 6. EBV data types

```{r}

df_data_type_counts <- compute_df_data.type(df = dataset)

df_data_type_counts

```

Plot the results

```{r}

plot_data_type <- plot_data.type_counts(df_data_type_counts)

#ggsave("plots_repo/plot_data_type.png", height = 6, width = 9)

```

### Combine plots

V1

```{r}

plot_combined_corpus <- ggarrange(
plot_temp_duration,
plot_spatial_range_counts,
plot_spatial_temporal_relevance,
plot_data_type,
nrow = 2,
ncol = 2,
  labels = c("A", "B", "C","D"),
heights = c(1,1.22))

plot_combined_corpus

#ggsave("plots_repo/plot_combined_corpus.png", height = 10, width = 12)
```

V2

```{r}

plot_combined_corpus.v2 <- ggarrange(
p_temp_range,
plot_spatial_range_counts,
plot_spatial_temporal_relevance,
plot_data_type,
nrow = 2,
ncol = 2,
  labels = c("A", "B", "C","D"),
heights = c(1,1.22))

plot_combined_corpus.v2

#ggsave("plots_repo/plot_combined_corpus.v2.png", height = 10, width = 12)
```


## 7. Data format


Those datasets in the repository that are relevance category X don't have format information.

```{r}

#plot_data_format_counts <- plot_data.type_format(dataset)

#plot_data_format_counts

#ggsave("plots_repo/plot_data_format_counts.png", height = 7, width = 9)


```






## 9. Journals


Count publications per journal

```{r}

dataset.j <- dataset[dataset$journal != "" & dataset$journal != "no",]

df_journals_counts <- as.data.frame(table(dataset.j$journal))

colnames(df_journals_counts) = c("Journal", "Publication_counts")

df_journals_counts <- df_journals_counts[order(-df_journals_counts$Publication_counts),]


df_journals_counts


```




```{r}

print(paste("A total of", length(df_journals_counts$Journal), "journals" ))

```



# Location of information - features



## % data type in abstract and title

```{r}
df_dataset_types <- read.csv("../data/dataset_types.csv", sep = ";")
df_dataset_types
```

I will add the following synonyms:

presence-absence: detection, capture

EBV genetic analyses: 16S, 18S, genetic data, microsatelites, barcodes(?), haplotypes, eDNA, SNPs


```{r}
df_dataset_types[nrow(df_dataset_types)+1,] <- c("presence-absence", "detection", "synonym")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("presence-absence", "capture", "synonym")

df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "16S", "methods")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "18S", "methods")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "barcodes", "methods")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "haplotypes", "methods")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "eDNA", "methods")
df_dataset_types[nrow(df_dataset_types)+1,] <- c("EBV genetic analyses", "SNPs", "methods")

df_dataset_types
```




```{r}
url <- c()
dataset_type <- c()
keyword_abs <- list()
keyword_title <- list()
#keyword_type <- c()
in_title <- c()
in_abstract <- c()



for (i in 1:nrow(dataset)) {
  
  
  
  dataset_type[i] <- dataset[i, "data_type"]
  url[i] <- dataset[i, "url"]
  
  
  
  
  # search on the abstract and note the list of keywords that match the dataset types

  keyword_abs[[i]] <- unique(get_keywords(input_string = dataset$full_text[i], 
               dataset_types =  df_dataset_types))
  
  # if the list of keywords has 1 or more entry, then note that we found information in the abstract
  
  if(length(keyword_abs[i][!sapply(keyword_abs[i], is.null)]) == 0){
    
    in_abstract[i] <- "no"
    
  } else if (length(keyword_abs[i][!sapply(keyword_abs[i], is.null)]) > 0){
    
    in_abstract[i] <- "yes"
    
  }
  
  
  

  # search on the title and note the list of keywords that match the dataset types

  keyword_title[[i]] <- unique(get_keywords(input_string = dataset$title[i], 
               dataset_types =  df_dataset_types))
  
  # if the list of keywords has 1 or more entry, then note that we found information in the abstract
  
  if(length(keyword_title[i][!sapply(keyword_title[i], is.null)]) == 0){
    
    in_title[i] <- "no"
    
  } else if (length(keyword_title[i][!sapply(keyword_title[i], is.null)]) > 0){
    
    in_title[i] <- "yes"
    
  } 
  
  
  
}


dataset$data.type_in_abstract <- in_abstract
dataset$data_type_in_title <- in_title



print(paste(length(which(in_abstract == "yes")), 
            "with dataset type or synonym explicit in the abstract, out of", length(in_abstract)))

print(paste("so", 
      round(length(which(in_abstract == "yes"))/length(in_abstract)*100, 1),
"%"))

print(paste(length(which(in_title == "yes")), 
            "with dataset type or synonym explicit in the title, out of", length(in_title)))

print(paste("so", 
      round(length(which(in_title == "yes"))/length(in_title)*100, 1),
"%"))



```






```{r}


df_features.l <- dataset[,c("temporal_range_position", "temporal_duration_position", "spatial_range_position", "species_location", "data.type_in_abstract", "data_type_in_title")]

# add id

df_features.l$id <- c(1:nrow(df_features.l))



# separate by commas

df_features.l1 <- df_features.l %>%               
  separate_rows(temporal_range_position, sep=",") %>% 
  separate_rows(temporal_duration_position, sep=",") %>% 
  separate_rows(spatial_range_position, sep=",") %>% 
  separate_rows(species_location, sep=",")



# Homogenize levels

levels(as.factor(df_features.l1$species_location))


df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == " source link")] <- "repository text"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == "source link")] <- "repository text"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == "source link abstract")] <- "repository text"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == " dataset")] <- "dataset"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == " source publication text")] <- "article"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == "source publication text")] <- "article"
df_features.l1$temporal_range_position[which(df_features.l1$temporal_range_position == "no")] <- "not given"



df_features.l1$temporal_duration_position[which(df_features.l1$temporal_duration_position == "source link")] <- "repository text"
df_features.l1$temporal_duration_position[which(df_features.l1$temporal_duration_position == "source link abstract")] <- "repository text"
df_features.l1$temporal_duration_position[which(df_features.l1$temporal_duration_position == "source publication text")] <- "article"
df_features.l1$temporal_duration_position[which(df_features.l1$temporal_duration_position == "no")] <- "not given"



df_features.l1$spatial_range_position[which(df_features.l1$spatial_range_position == "source link")] <- "repository text"
df_features.l1$spatial_range_position[which(df_features.l1$spatial_range_position == "source link abstract")] <- "repository text"
df_features.l1$spatial_range_position[which(df_features.l1$spatial_range_position == "source publication text")] <- "article"
df_features.l1$spatial_range_position[which(df_features.l1$spatial_range_position == "no")] <- "not given"



df_features.l1$species_location[which(df_features.l1$species_location == "source link abstract")] <- "repository text"
df_features.l1$species_location[which(df_features.l1$species_location == "source publication text")] <- "article"
df_features.l1$species_location[which(df_features.l1$species_location == " source publication text")] <- "article"


df_features.l1

```

How many with all the features readily accessible (in repository text):

```{r}
n_all <- nrow(na.omit(df_features.l1[ 
                   df_features.l1$data.type_in_abstract == "yes" & 
                   df_features.l1$temporal_range_position == 'repository text' &
                   df_features.l1$spatial_range_position == 'repository text',]))

print(paste(n_all, "with all the features readily accessible (in the repository text) - a", round(n_all/nrow(dataset)*100, 3),"%"))

```

How many with at least one feature accessible

```{r}

df_features.l2 <- df_features.l1[!is.na(df_features.l1$data.type_in_abstract),]

df_features.l2 <- df_features.l1[ 
  df_features.l1$data_type_in_title == "yes" |
                   df_features.l1$data.type_in_abstract == "yes" |
                   df_features.l1$temporal_range_position == 'repository text' |
                   df_features.l1$spatial_range_position == 'repository text', ]

df_features.l2[!is.na(df_features.l2$data.type_in_abstract),]

print(paste(n_any, "with at least one feature readily accessible (in the repository text) - a", round(n_any/nrow(dataset)*100, 3),"%"))

```


How many per feature type



```{r}

# temporal range

dataset$id <- c(1:nrow(dataset))

tempr_position <- dataset[,c("temporal_range_position","id")]

tempr_position <- na.omit(tempr_position)


print(paste(nrow(tempr_position), "analyzed for temp range position"))

tempr_position <- tempr_position %>%               
  separate_rows(temporal_range_position, sep=",") 


tempr_position[which(tempr_position$temporal_range_position == " dataset"),"temporal_range_position"] <- "dataset"

tempr_position[which(tempr_position$temporal_range_position == " source publication text"),"temporal_range_position"] <- "article"

tempr_position[which(tempr_position$temporal_range_position == "source publication text"),"temporal_range_position"] <- "article"

tempr_position[which(tempr_position$temporal_range_position == " source link"),"temporal_range_position"] <- "repository text"

tempr_position[which(tempr_position$temporal_range_position == "source link"),"temporal_range_position"] <- "repository text"

tempr_position[which(tempr_position$temporal_range_position == "source link abstract"),"temporal_range_position"] <- "repository text"

tempr_position[which(tempr_position$temporal_range_position == "no"),"temporal_range_position"] <- "not given"

tempr_position



# temporal duration

dataset$id <- c(1:nrow(dataset))

tempd_position <- dataset[,c("temporal_duration_position","id")]

tempd_position <- na.omit(tempd_position)


print(paste(nrow(tempd_position), "analyzed for temp range position"))

tempd_position <- tempd_position %>%               
  separate_rows(temporal_duration_position, sep=",") 


tempd_position[which(tempd_position$temporal_duration_position == " dataset"),"temporal_duration_position"] <- "dataset"

tempd_position[which(tempd_position$temporal_duration_position == " source publication text"),"temporal_duration_position"] <- "article"

tempd_position[which(tempd_position$temporal_duration_position == "source publication text"),"temporal_duration_position"] <- "article"

tempd_position[which(tempd_position$temporal_duration_position == " source link"),"temporal_duration_position"] <- "repository text"

tempd_position[which(tempd_position$temporal_duration_position == "source link"),"temporal_duration_position"] <- "repository text"

tempd_position[which(tempd_position$temporal_duration_position == "source link abstract"),"temporal_duration_position"] <- "repository text"

tempd_position[which(tempd_position$temporal_duration_position == "no"),"temporal_duration_position"] <- "not given"

tempd_position


# spatial range duration

dataset$id <- c(1:nrow(dataset))

spatialr_position <- dataset[,c("spatial_range_position","id")]

spatialr_position <- na.omit(spatialr_position)


print(paste(nrow(spatialr_position), "analyzed for temp range position"))

spatialr_position <- spatialr_position %>%               
  separate_rows(spatial_range_position, sep=",") 


spatialr_position[which(spatialr_position$spatial_range_position == " dataset"),"spatial_range_position"] <- "dataset"

spatialr_position[which(spatialr_position$spatial_range_position == " source publication text"),"spatial_range_position"] <- "article"

spatialr_position[which(spatialr_position$spatial_range_position == "source publication text"),"spatial_range_position"] <- "article"

spatialr_position[which(spatialr_position$spatial_range_position == " source link"),"spatial_range_position"] <- "repository text"

spatialr_position[which(spatialr_position$spatial_range_position == "source link"),"spatial_range_position"] <- "repository text"

spatialr_position[which(spatialr_position$spatial_range_position == "source link abstract"),"spatial_range_position"] <- "repository text"

spatialr_position[which(spatialr_position$spatial_range_position == "no"),"spatial_range_position"] <- "not given"


 



df_tempr <- as.data.frame(table(tempr_position$temporal_range_position))
df_tempr$feature <- rep("temporal range", times = nrow(df_tempr))

df_tempd <- as.data.frame(table(tempd_position$temporal_duration_position))
df_tempd$feature <- rep("temporal duration", times = nrow(df_tempd))

df_spatr <- as.data.frame(table(spatialr_position$spatial_range_position))
df_spatr$feature <- rep("spatial range", times = nrow(df_spatr))

df_locations_plot <- rbind(df_tempr, df_tempd, df_spatr)

colnames(df_locations_plot) <- c("location", "Freq", "feature")

df_locations_plot$location <- factor(df_locations_plot$location, levels = c("not given", "article","dataset", "repository text"))


ggplot(df_locations_plot, aes(x=location, y = Freq, group = feature)) +
   geom_segment( aes(x=location ,xend=location, y=0, yend=max(Freq)), color="grey") +
    geom_point(aes(shape = feature),size=4, alpha = 0.8) +
    coord_flip() +
  theme_bw() +
  my.theme+
  #scale_color_manual(values = c("deepskyblue2","firebrick1","indianred4"))+
  scale_shape_manual(values = c(16,8,2))+
  ylab("publication counts")

#ggsave("plots_repo/location_features.png", height = 5, width = 9)

```






## Spatial information format


```{r}


geospatial_info_article_text <- dataset$geospatial_info_article_text[-which(is.na(dataset$geospatial_info_article_text))]

geospatial_info_dataset <- dataset$geospatial_info_dataset[-which(is.na(dataset$geospatial_info_dataset))]

geospatial_info_repo_page_text <- dataset$geospatial_info_repo_page_text[-which(is.na(dataset$geospatial_info_repo_page_text))]






geospatial_info_article_text  <- unlist(strsplit(geospatial_info_article_text,","))

geospatial_info_dataset  <- unlist(strsplit(geospatial_info_dataset,","))

geospatial_info_repo_page_text  <- unlist(strsplit(geospatial_info_repo_page_text,","))


geospatial_info_article_text[which(geospatial_info_article_text == " administrative unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == " range coordites")] <- "range coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == " site coordites")] <- "site coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == "site coordites")] <- "site coordinates"
geospatial_info_article_text[which(geospatial_info_article_text == " geological feature me")] <- "geographic feature"
geospatial_info_article_text[which(geospatial_info_article_text == "geological feature me")] <- "geographic feature"
geospatial_info_article_text[which(geospatial_info_article_text == " Map")] <- "Map"
geospatial_info_article_text[which(geospatial_info_article_text == "Map")] <- "map"
geospatial_info_article_text[which(geospatial_info_article_text == "administration unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == "Administrative unit")] <- "administrative unit"
geospatial_info_article_text[which(geospatial_info_article_text == "no")] <- "not given"


geospatial_info_dataset[which(geospatial_info_dataset == " site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "  site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "site coordites")] <- "site coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == "sample coordites")] <- "sample coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == " sample coordites")] <- "sample coordinates"
geospatial_info_dataset[which(geospatial_info_dataset == " administrative unit")] <- "administrative unit"
geospatial_info_dataset[which(geospatial_info_dataset == " site IDs")] <- "site IDs"
geospatial_info_dataset[which(geospatial_info_dataset == " IDs")] <- "site IDs"
geospatial_info_dataset[which(geospatial_info_dataset == " geographic feature")] <- "geographic feature"
geospatial_info_dataset[which(geospatial_info_dataset == "no")] <- "not given"


geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "geological feature me")] <- "geographic feature"
geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "Administrative unit")] <- "administrative unit"
geospatial_info_repo_page_text[which(geospatial_info_repo_page_text == "no")] <- "not given"

```


```{r}

location <- rep("article text", times = length(geospatial_info_article_text))

df_article <- data.frame(geospatial_info_article_text, location)

location <- rep("repository page", times = length(geospatial_info_repo_page_text))

df_repo<- data.frame(geospatial_info_repo_page_text, location)

location <- rep("dataset", times = length(geospatial_info_dataset))

df_dataset <- data.frame(geospatial_info_dataset, location)

colnames(df_article) <- c("information", "location")
colnames(df_repo) <- c("information", "location")
colnames(df_dataset) <- c("information", "location")

df_locations <- rbind(df_article,df_repo,df_dataset)

df_locations$information[which(df_locations$information == "site IDs") ] <- "sites IDs"
df_locations$information[which(df_locations$information == "site coordites") ] <- "site coordinates"
df_locations$information[which(df_locations$information == "Map") ] <- "map"
df_locations$information[which(df_locations$information == "geologial feature") ] <- "geographic feature"
df_locations$information[which(df_locations$information == "geographical feature") ] <- "geographic feature"

```


```{r}

df_locations <- as.data.frame(table(df_locations))


df_locations$location <- factor(df_locations$location, levels=c("article text", "repository page", "dataset"))

df_locations$information <- factor(df_locations$information, levels=c("not given", "sites IDs", "map", "administrative unit", "geographic feature", "distribution model", "range coordinates", "site coordinates", "sample coordinates"))



ggplot(df_locations, aes(x=information, y = Freq, group = location, color = location)) +
   geom_segment( aes(x=information ,xend=information, y=0, yend=max(Freq)), color="grey") +
    geom_point(size=4, alpha = 0.5) +
    coord_flip() +
  theme_bw() +
  my.theme+
 scale_color_manual(values = c("blue4","red1","green4"))+
 # scale_shape_manual(values = c(11,2,3))+
  ylab("publication counts")+
  xlab("geospatial information")

  
#ggsave("plots_repo/location_type_geoinfo.png", height = 5, width = 9)


```



```{r}
length(dataset$data.type_in_abstract)

dataset[dataset$data.type_in_abstract == "no" & dataset$data_type_in_title == "no",]

length(dataset$data_type_in_title[dataset$data_type_in_title == "yes"])
length(dataset$data.type_in_abstract[dataset$data.type_in_abstract == "yes"])
```






